---
title: "segregation_measures"
author: "Rebecca Amodeo"
date: "4/10/2020"
## Oregon's Racial Segregation

1.	Residential segregation in Oregon: <https://www.oregon.gov/OHA/PH/ABOUT/Documents/indicators/segregation.pdf>
2.	Desegregation and Multiculturalism in the Portland Public Schools: <https://www.ohs.org/research-and-library/oregon-historical-quarterly/upload/OHQ_111_1-Johnson-and-Williams_PPS.pdf>
3.	Timeline of Oregon and U.S. Racial, Immigration, and Educational History: <https://www.portlandoregon.gov/bps/article/412697>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(sp)
library(sf)
library(seg)
```

## Segregation Measures
Modifying Daniel's segregation script for my indicators.

### Safety Index
```{r safety, echo = FALSE}





```


### Economics Indicators
```{r Seg, echo = FALSE}
geo_tracts <- tigris::tracts("OR")

tracts <- st_as_sf(geo_tracts) %>% 
  select(GEOID) %>% 
  left_join(or_wide) %>%
  mutate(NAME = gsub("^(.+\\d),\\s.+", "\\1", NAME))

calc_seg <- function(data, var1, var2, bw = c(0.3, 0.7)) {
  mods <- map(bw,
    ~spseg(as_Spatial(data$geometry), 
      select(st_drop_geometry(data), 
        !!enquo(var1), 
        !!enquo(var2)), 
      smoothing = "kernel", 
      sigma = .x)
  )
  names(mods) <- bw
  mods
}

pull_seg_estimates <- function(l) {
  tibble(binwidth = names(l),
         d = map_dbl(l, ~.x@d),
         r = map_dbl(l, ~.x@r),
         h = map_dbl(l, ~.x@h))
}

tracts_seg <- tracts %>%
  nest(data = c(NAME, pre_cap_wht_inc_trc, per_cap_inc_trc, geometry)) %>%
  mutate(seg_mod = map(data, calc_seg, pre_cap_wht_inc_trc, per_cap_inc_trc, bw = 0.3), seg_est = map(seg_mod, pull_seg_estimates)) %>%
  select(county, seg_est) %>%
  unnest(seg_est)

write_csv(tracts_seg, here::here("data", "or-county-seg-whtinc.csv"))

#tracts_seg <- read_csv(here::here("data", "or-county-seg.csv"))

counties <- tigris::counties("OR", cb = TRUE) %>%
  st_as_sf() %>%
  mutate(county = paste(NAME, "County"))

left_join(counties, tracts_seg)  %>%
  mapview::mapview(zcol = "d")

districts <- tigris::school_districts("OR") %>%
  st_as_sf()

#Just a bit of extra code if you want a quick interactive map. Note
#there’s… not a lot of between-county variance in these estimates.


counties <- tigris::counties("OR", cb = TRUE) %>%
  st_as_sf() %>%
  mutate(county = paste(NAME, "County"))

left_join(counties, tracts_seg)  %>%
  mapview::mapview(zcol = "d")

```